Classless Java: Tuning Java Interfaces

### 2015-11-27 paper/sections/RelatedWork.tex
more related work

\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
strong line between units of reuse (traits) and object factories
(classes). In this model, traits as units of reusable code, contain
only methods as reusable entities. Thus, no state and state
initialization are considered. Classes act as object factories,
requiring functionalities from multiple traits. Java 8 interfaces with
default methods are closely related to traits: concrete method
implementation are allowed (via the \textbf{default} keyword) inside
interfaces. The introduction of default methods opens the gate for
various flavours of multiple inheritance in Java, using interfaces.
Traits offer a trait algebra with operations like sum, alias and exclusion, provided
for explicit conflict resolution.  
Former work by Bono et al.~\cite{bono14}. provides details on
mimicking the trait algebra through Java 8 interfaces. 
We briefly recall the main points of their encoding;
however we propose a different representation of \textbf{exclusion}.
The author of ~\cite{bono14} agree that our revised version is
cleaner, typesafe and more direct.\bruno{do we want do say this?}\yanlin{need to
  ask Marco.}

@@

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.
In the trait model two concepts (traits and classes) coexist and
cooperate. Some authors~\bruno{who?}\yanlin{need to ask Marco} see this as good language design
fostering good software development by helping programmers to think
about the structure of their programs.  However, other authors see the
need of two concepts and the absence of state as drawbacks of this
model~\cite{malayeri2009cz}.  Object interfaces are units of reuse, and at
the same time they provide factory methods for instantiation and
support state.  Our approach promotes the use of interfaces instead of
classes, in order to rely on the modular composition offered
by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation
driven code generation is enough to overcome this difficulty, and the
resulting programming style encourages modularity, composability and
reusability, by keeping a strong object oriented feel. In that sense,
we promote object interfaces as being both units of reusable code and
object factories. Our practical experience is that, in Java,
separating the two notions leads to alot of boilerplate code, and is
quite limiting when multiple inheritance with state is required. The
use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite expressive. 
Moreover the ability to support constructors 
adds additional expressivity, which is not available in approaches such 
as Scala's traits/mixins.

@@

One famous attempt in this direction is provided by
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
\textbf{this}, changing its meaning along with inheritance. However when
invoking a method with MyType in a parameter position, the exact type of the
receiver must be known.  This is a big limitation in class based object oriented
programming, and is exasperated by the interface-based programming we propose: no
type is ever going to be exact since classes are not explicitly used. A recent
article~\cite{Saito2013933} lights up this topic, proposing two
new features: exact statements and nonheritable methods. Both are
related to our work: 1) any method generated inside of the \Q@of@ method is indeed
non-inheritable since there is no class name to extend from; 2)exact
statements (a form of wild-card capture on the exact run-time type) could
capture the ``exact type'' of an object even in a class-less environment.
%Interestingly, our use of covariant type refinement is enough to emulate

% The addition of MyType to a language will allow easy definition of
%binary methods, methods with recursive types (i.e., the same type of the
%receiver appears in the argument or return positions of methods), etc. MyType
%greatly enhances the expressiveness and extensibility of object-oriented
%programming languages.

%In our approach, we are using covariant-return types to simulate some uses of
%MyType. But our approach only works on method positive positions, whereas MyType
%is more general, as it works on any positions. Nevertheless our approach is
%still useful for modeling fluent interfaces and solving expression
%problems,etc.

@@

Here (compile/load time)\yanlin{only compile time should do?} code generation comes at the rescue: if for a certain
feature (\mixin in our case) it is possible to use the original language syntax
to \emph{express-describe} any specific instantiation of such feature
(annotating a class and provide getters), then we can insert in the compilation
process a tool that examins and enrichs the code before compilation. No need to
modify the original source; for example we can work on temporary files
\yanlin{Here what does temporary files mean? could explain more?}. Java is
a particular good candidate for this kind of manipulation since it already
provide ways to define and integrate such tools in its own compilation process:
in this way there is no need of temporary files, and there is a well defined way
of putting multiple extensions together.

@@

\subsection{Formalization of Java8}
We provide a simple formalization for a subset of Java
including default/static interface methods and object initialization literals
(often called anonymous local inner classes).  A similar formalization was
drafted by Goetz and Field~\cite{goetz12fdefenders} to formalize defender
(default) methods in Java. However this formalization is limited to model
exactly one method inside classes/interfaces.\bruno{explain why having multiple 
methods is significantly more challenging.}

>>>>

\paragraph{Traits and Java's default methods}
Simplifying the mixins approach, traits~\cite{scharli03traits} draw a
strong line between units of reuse (traits) and object factories
(classes). In this model, traits as units of reusable code, contain
only methods as reusable entities. Thus, no state and state
initialization are considered. Classes act as object factories,
requiring functionalities from multiple traits. Java 8 interfaces with
default methods are closely related to traits: concrete method
implementation are allowed (via the \textbf{default} keyword) inside
interfaces. The introduction of default methods opens the gate for
various flavours of multiple inheritance in Java, using interfaces.
Traits offer a trait algebra with operations like sum, alias and exclusion, provided
for explicit conflict resolution.  
Former work by Bono et al.~\cite{bono14}. provides details on
mimicking the trait algebra through Java 8 interfaces. 
We briefly recall the main points of their encoding;
however we propose a different representation of \textbf{exclusion}.
The author of ~\cite{bono14} agree that our revised version is
cleaner, typesafe and more direct.
%\bruno{do we want do say this?}\yanlin{need to ask Marco.}

@@

\paragraph{Traits vs Object Interfaces}
We consider object interfaces to be an alternative to traits or mixins.
In the trait model two concepts (traits and classes) coexist and
cooperate. Some authors~\cite{bettini2013pure}\yanlin{There are many trait related papers of ferruccio
  damiani, please check whether citing this one is ok.} 
see this as good language design
fostering good software development by helping programmers to think
about the structure of their programs.  However, other authors see the
need of two concepts and the absence of state as drawbacks of this
model~\cite{malayeri2009cz}.  Object interfaces are units of reuse, and at
the same time they provide factory methods for instantiation and
support state.  Our approach promotes the use of interfaces instead of
classes, in order to rely on the modular composition offered
by interfaces. Since Java was designed for classes, a direct class-less
programming style is verbose and feels unnatural. However, annotation
driven code generation is enough to overcome this difficulty, and the
resulting programming style encourages modularity, composability and
reusability, by keeping a strong object oriented feel. In that sense,
we promote object interfaces as being both units of reusable code and
object factories. Our practical experience is that, in Java,
separating the two notions leads to alot of boilerplate code, and is
quite limiting when multiple inheritance with state is required. The
use of abstract state operations avoids the key difficulties
associated with multiple inheritance and state, while still being quite expressive. 
Moreover the ability to support constructors 
adds additional expressivity, which is not available in approaches such 
as Scala's traits/mixins.

@@

One famous attempt in this direction is provided by
\emph{MyType}~\cite{bruce1994paradigmatic}, representing the type of
\textbf{this}, changing its meaning along with inheritance. However when
invoking a method with MyType in a parameter position, the exact type of the
receiver must be known.  This is a big limitation in class based object oriented
programming, and is exasperated by the interface-based programming we propose: no
type is ever going to be exact since classes are not explicitly used. A recent
article~\cite{Saito2013933} lights up this topic, proposing two
new features: exact statements and nonheritable methods. Both are
related to our work: 1) any method generated inside of the \Q@of@ method is indeed
non-inheritable since there is no class name to extend from; 2)exact
statements (a form of wild-card capture on the exact run-time type) could
capture the ``exact type'' of an object even in a class-less
environment. 
Admittedly, MyType greatly enhances the expressiveness and extensibility of
object-oriented programming languages. We are using covariant-return types to
simulate some uses of MyType. But our approach only works on method positive
positions, whereas MyType is more general, as it works on any
positions. Nevertheless our approach is still useful for modeling fluent
interfaces and solving expression problems,etc.
% The addition of MyType to a language will allow easy definition of
%binary methods, methods with recursive types (i.e., the same type of the
%receiver appears in the argument or return positions of methods), etc.

@@

Here compile time code generation comes at the rescue: if for a certain feature
(\mixin in our case) it is possible to use the original language syntax to
\emph{express-describe} any specific instantiation of such feature (annotating a
class and provide getters), then we can insert in the compilation process a tool
that examins and enrichs the code before compilation. No need to modify the
original source; for example we can work on temporary files Java is a particular
good candidate for this kind of manipulation since it already provide ways to
define and integrate such tools in its own compilation process: in this way
there is no need of temporary files, and there is a well defined way of putting
multiple extensions together.

@@

\subsection{Formalization of Java8}
We provide a simple formalization for a subset of Java including default/static
interface methods and object initialization literals (often called anonymous
local inner classes).  A similar formalization was drafted by Goetz and
Field~\cite{goetz12fdefenders} to formalize defender (default) methods in
Java. However this formalization is limited to model exactly one method inside
classes/interfaces.\bruno{explain why having multiple methods is significantly
  more challenging.} Our code generation generates several methods that may
interacts with each other and methods defined by user. Thus implementing and
formalizing multiple methods inside interfaces adds complexity greatly.
